---
title: Chi-Square Difference Testing Using the Satorra-Bentler Scaled Chi-Square with
  Results from Mplus
author: Xidan Chen
date: '2018-11-07'
slug: chi-square-difference-testing-using-the-satorra-bentler-scaled-chi-square-with-results-from-mplus
categories: []
tags:
  - Mplus
  - R
  - multiple group comparison
  - SEM
  - chi-square difference test
  - nested models
---
I was working on a project for which I need to run around 50 sets of comparisons of nested models. For the non-normal continuous data I was working with, I used the MLR as estimation method. Mplus unfortunately won’t compute the chi-square difference test using MLR chi-square. Following their guide here http://www.statmodel.com/chidiff.shtml, I therefore wrote these functions to automate my analysis. Although these functions were written for my own analysis using MLR, the same functions could also be used for MLM (Satorra-Bentler) and WLSM chi-square. Below is the guide. 

Required packages are **dplyr, MplusAutomation, plyr**. 

1. Load these functions into your R: **outFile, SumChiSq, diffTestMLR**
--
```{r}

outFile <- function(test.dir, models = c("lessRestrictive", "restrictive"), list.of.files){
  dir.create(test.dir)
  new.folder <- paste0(test.dir, models, collapse='')
  dir.create(new.folder)
  file.copy(list.of.files, new.folder)
  
}

SumChiSq <- function(diff.dir){
  library(plyr)
  library(dplyr)
  library(MplusAutomation)
  
  allOutput <- readModels(diff.dir, recursive=TRUE)
  justSummaries <- do.call("rbind.fill",sapply(allOutput,"[", "summaries"))
  justSummaries %>%
         select(Filename, 
         #Observations, 
         ChiSqM_Value, 
         ChiSqM_DF, ChiSqM_ScalingCorrection)
  
}


diffTestMLR <- function(lessRestrictive.dir, restrictive.dir){
  
  lessRestrictive <- SumChiSq(lessRestrictive.dir)
  restrictive <- SumChiSq(restrictive.dir)
#Scaled χ2 = (F0c0 - F1c1)(d0 - d1)/(c0d0 - c1d1)
# F0 = The χ2 value from the constrained model.
# F1 = The χ2 value from the freely estimated model.
# c0 = The scaling factor from the constrained model.
# c1 = The scaling factor from the freely estimated model.
# d0 = The degrees of freedom from the constrained model.
# d1 = The degrees of freedom from the freely estimated model.

old_names <- c("Filename", "ChiSqM_Value", "ChiSqM_DF", "ChiSqM_ScalingCorrection")
new_namesf0 <- c("Filename", "F0", "d0", "c0")
new_namesf1 <- c("Filename", "F1", "d1", "c1")

restrictive2 <- restrictive %>% 
  rename_at(vars(old_names), ~ new_namesf0)
lessRestrictive2 <- lessRestrictive %>% 
  rename_at(vars(old_names), ~ new_namesf1)

sumChiSq2 <- merge(restrictive2, lessRestrictive2， by="Filename")

sumChiSq2 <- sumChiSq2 %>% 
  mutate(df = d0 - d1) %>%
  mutate(ScaledChiSq = (F0*c0 - F1*c1)*(d0 - d1)/(c0*d0 - c1*d1)) 

sumChiSq2$p_value <- pchisq(sumChiSq2$ScaledChiSq, sumChiSq2$df, lower.tail=FALSE)
print(list('lessRestrictive:' = lessRestrictive, 'restrictive:' = restrictive, 'diffTestMLR:' = sumChiSq2))
  
}


```



2. Example
--
What you need to do first is specifying a test directory, the list of files for your nested models, and model type (restrictive or less restrictive). **outFile** will create a directory for the difftest, copy the Mplus outputs to the specific model type directories you choose. 
```{r, eval=FALSE}

test.dir <- './diffTest/'
list.of.files1 <- c("./MplusCode/STEP1/tap3/tap3.out", "./MplusCode/STEP1/vlmta3/vlmta3.out")
outFile(test.dir, "lessRestrictive", list.of.files1)

test.dir <- './diffTest/'
list.of.files2 <- c("./MplusCode/STEP2/tap3/tap3.out", "./MplusCode/STEP2/vlmta3/vlmta3.out")
outFile(test.dir, "restrictive", list.of.files2)
```

**SumChiSq will** output the chi-square statistics under the diff test directory

**diffTestMLR** will output the diff test results when restrictive models and less restrictive models are specified. You could skip the **outFile**, use **diffTestMLR** function directly if you don't have many comparisons.

For example:
```{r, warning=FALSE, message=FALSE}
lessRestrictive.dir <- './diffTest/lessRestrictive'
restrictive.dir <- './diffTest/Restrictive'
diffTestMLR(lessRestrictive.dir, restrictive.dir)
```






